# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
data <- read.csv("activity.csv", header=TRUE,
                 colClasses=c("integer", "Date", "integer"))
```

## What is mean total number of steps taken per day?

```{r}
library(lattice)
steps_per_day <- aggregate(steps ~ date, data, FUN = sum)
histogram(~ steps, steps_per_day,
          breaks = 15,
          main = "Distribution of the total number of steps taken each day",
          xlab = "Number of steps per day",
          ylab = "Percent")
```

```{r}
mean_steps_per_day <- mean(steps_per_day$steps)
median_steps_per_day <- median(steps_per_day$steps)
```

Mean is **`r mean_steps_per_day`** steps per day. 
Median is **`r median_steps_per_day`** steps per day.

## What is the average daily activity pattern?

```{r}
steps_across_days <- aggregate(steps ~ interval, data, FUN = mean)
xyplot(steps ~ interval, data=steps_across_days, type='l',
       main = "Number of steps per 5 min averaged across days",
       xlab = "Time of the day",
       ylab = "Steps")
```

```{r}
max_interval <- steps_across_days$interval[which.max(steps_across_days$steps)]
```

Five minite interval with the maximum number of steps is **`r max_interval`**.

## Imputing missing values

```{r}
total_na <- sum(is.na(data$steps))
```

The total number of rows with NA is **`r total_na`**. 

Missing values are imputed with the average steps for the corresponding interval.

```{r}
steps_imputed <- apply(data, 1,
                       function(x) if (is.na(as.integer(x[1])))
                         as.integer(steps_across_days$steps[which(steps_across_days$interval == as.integer(x[3]))])
                       else
                         as.integer(x[1]))
data_imputed <- cbind(steps_imputed, data[, 2:3])
```

```{r}
steps_per_day_imputed <- aggregate(steps_imputed ~ date, data_imputed, FUN = sum)
histogram(~ steps_imputed, steps_per_day_imputed,
          breaks = 15,
          main = "Distribution of the total number of steps taken each day (imputed)",
          xlab = "Number of steps per day (imputed)",
          ylab = "Percent")
```

```{r}
mean_steps_per_day_imputed <- mean(steps_per_day_imputed$steps_imputed)
median_steps_per_day_imputed <- median(steps_per_day_imputed$steps_imputed)
```

Mean is **`r mean_steps_per_day`** steps per day. 
Median is **`r median_steps_per_day`** steps per day.

```{r}
total_steps_per_day <- sum(steps_per_day$steps)
total_steps_per_day_imputed <- sum(steps_per_day_imputed$steps_imputed)
```

Total daily number of steps is **`r total_steps_per_day`** steps. 
Total daily number of steps after imputation is **`r total_steps_per_day_imputed`** steps. 

The imputation has **little impact** on the mean value, even though it **increases** the total value **significantly**.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
Sys.setlocale("LC_TIME", "C")
```

```{r}
data <- cbind(data, sapply(data$date,
                           function(x) if (weekdays(x) ==
                              "Saturday" | weekdays(x) == "Sunday")
                             "Weekend"
                           else
                             "Weekday"
                             )
              )
names(data)[4] <- "weekend"

xyplot(steps ~ interval | weekend, data=data, layout=c(1, 2), type='l')
```
